---
title: "p8105_hw3_fm2751"
output: github_document
date: "2022-10-15"
---

```{r setup, include=FALSE}
library(tidyverse)
library(ggridges)
library(patchwork)

library(p8105.datasets)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```

##  Problem 0
Public GitHub repo, local R Project, subdirectory for data, and .Rmd file created and rendered to github_document.

## Problem 1

### Read in the data

```{r}
data("instacart")

instacart = 
  instacart %>% 
  as_tibble(instacart)
```

### Answer questions about the data

This dataset contains `r nrow(instacart)` rows and `r ncol(instacart)` columns, with each row resprenting a single product from an instacart order. Variables include identifiers for user, order, and product; the order in which each product was added to the cart. There are several order-level variables, describing the day and time of the order, and number of days since prior order. Then there are several item-specific variables, describing the product name (e.g. Yogurt, Avocado), department (e.g. dairy and eggs, produce), and aisle (e.g. yogurt, fresh fruits), and whether the item has been ordered by this user in the past. In total, there are `r instacart %>% select(product_id) %>% distinct %>% count` products found in `r instacart %>% select(user_id, order_id) %>% distinct %>% count` orders from `r instacart %>% select(user_id) %>% distinct %>% count` distinct users.

Below is a table summarizing the number of items ordered from aisle. In total, there are 134 aisles, with fresh vegetables and fresh fruits holding the most items ordered by far.

```{r}
  instacart %>% 
    count(aisle) %>% 
  arrange(desc(n))
```

Next is a plot that shows the number of items ordered in each aisle. Here, aisles are ordered by ascending number of items.

```{r}
instacart %>% 
  count(aisle) %>% 
  filter(n > 10000) %>% 
  mutate(aisle = fct_reorder(aisle, n)) %>% 
  ggplot(aes(x = aisle, y = n)) + 
  geom_point() + 
  labs(title = "Number of items ordered in each aisle") +
  theme(axis.text.x = element_text(angle = 60, hjust = 1))
```

Our next table shows the three most popular items in aisles `baking ingredients`, `dog food care`, and `packaged vegetables fruits`, and includes the number of times each item is ordered in your table.
* In `packaged vegetables fruits`, the three most popular items are Organic Baby Spinach (9784), Organic Raspberries (5546), and Organic Blueberries (4966);
* In `baking ingredients`, the three most popular items are Light Brown Sugar (499), Pure Baking Soda (387), and Cane Sugar (336);
* In `dog food care`, the three most popular items are Snack Sticks Chicken & Rice Recipe Dog Treats (30), Organix Chicken & Brown Rice Recipe (28), and Small Dog Biscuits (26).

```{r}
instacart %>% 
  filter(aisle %in% c("baking ingredients", "dog food care", "packaged vegetables fruits")) %>%
  group_by(aisle) %>% 
  count(product_name) %>% 
  mutate(rank = min_rank(desc(n))) %>% 
  filter(rank < 4) %>% 
  arrange(desc(n)) %>%
  knitr::kable()
```

Finally is a table showing the mean hour of the day at which Pink Lady Apples and Coffee Ice Cream are ordered on each day of the week. This table has been formatted in an untidy manner for human readers. Pink Lady Apples are generally purchased slightly earlier in the day than Coffee Ice Cream, with the exception of day 5.

```{r}
instacart %>%
  filter(product_name %in% c("Pink Lady Apples", "Coffee Ice Cream")) %>%
  group_by(product_name, order_dow) %>%
  summarize(mean_hour = mean(order_hour_of_day)) %>%
  spread(key = order_dow, value = mean_hour) %>%
  knitr::kable(digits = 2)
```

## Problem 2

### Load and tidy the data
```{r}
accel = 
  read_csv("./data/accel_data.csv", col_types = cols()) %>% 
  janitor::clean_names() %>% 
  pivot_longer(
    activity_1:activity_1440,
    names_prefix = "activity_",
    names_to = "time_in_mins",
    values_to = "activity_counts") %>% 
  mutate(
    time_in_mins = as.numeric(time_in_mins),
    week = as.factor(week),
    day = factor(day, levels = c("Sunday", "Monday", "Tuesday", "Wednesday", "Thursday", "Friday", "Saturday")),
    weekday_vs_weekend  = ifelse(day %in% c("Sunday", "Saturday"), print("weekend"), print("weekday"))) %>%
  arrange(day)
```

#### Description of the resulting dataset
The resulting dataset `accel` contains `r nrow(accel)` observations and `r ncol(accel)` variables, they are: `r names(accel)`. It's worth pointing out that activity time is recorded in minutes, and we added a new variable `weekday_vs_weekend` to the original dataset to indicate whether the day is a weekday or a weekend, which can be very helpful in the following analyses.

### Generate a table of total activity/day
```{r}
tot_act <-
  accel %>% 
  group_by(day_id) %>% 
  summarise(total_activity = sum(activity_counts))
tot_act %>% knitr::kable()
```

### Generate a plot of total activity/day to see if any trend present
```{r}
tot_act %>% 
  ggplot(aes(x = day_id, y = total_activity)) +
  geom_point() +
  geom_line() +
  labs (
    title = "Total activity over each day",
    x = "Day_id",
    y = "Total activity"
  ) +
  theme_classic()
```
* According to the above table and plot generated from the `accel` dataset, when we focus on the total activity for each day (i.e., `total_activity` vs. `day_id`), there is no any apparent trends observed. We might need to focus on different groupings to observe an apparent trend.




